A Survey of Current Exoskeletons\\ and Their Control Architectures and Algorithms\\ (Draft 4.0)

### 2015-10-01 exos/bleex.tex
more updates

The BLEEX team has successfully implemented both sensitivity amplification \cite{sesitivityAmpPaper2005} and a hybrid assitive control scheme \cite{bleex_hybrid_control_2006} that switches (based on gait phase) between sensitivity amplification and a position control regulating desired torque. This section focuses on sensitivity amplification, as the hybrid assitive strategy did not perform as well in walking trials.

@@

In its sensitivity amplification implementation, BLEEX controllers attempt to minimize interaction forces between the human pilot and the exoskeleton. However, the BLEEX team makes a point to avoid direct measurements of the interaction force between the human and the exoskeleton suit.  They highlight difficulties in properly outfitting humans with necessary sensing equipment and challenges in modeling the interaction between the human and exoskeleton (e.g., non-rigid, non-fixed contact points).  Instead, the team develops controllers based on an inverse dynamic model (in the saggital plane) of only the exoskeleton.

@@

The BLEEX sensitivity amplification control strategy models the torque applied by the human pilot on the exoskeleton as ${\bf d}$ (assuming no outside disturbances).   Neglecting gravity, the exokeleton angular velocity is modeled as 
\[\vv = \mG \vr + \mS {\bf d} ,\] 
where $\mG$ is the transfer function from actuator inputs ($\mG$ is the exoskeleton's dynamics), $\vr$ is the actuator input, and $\mS$ is the sensitivity or transfer function from human torque to exokeleton angular velocity.

>>>>

The BLEEX team has successfully implemented both sensitivity amplification \cite{sesitivityAmpPaper2005} and a hybrid assistive control scheme \cite{bleex_hybrid_control_2006} that switches (based on gait phase) between sensitivity amplification and a position control regulating desired torque. This section focuses on sensitivity amplification, as the hybrid assistive strategy did not perform as well in walking trials.

@@

In its sensitivity amplification implementation, BLEEX controllers attempt to minimize interaction forces between the human pilot and the exoskeleton. However, the BLEEX team makes a point to avoid direct measurements of the interaction force between the human and the exoskeleton suit.  They highlight difficulties in properly outfitting humans with necessary sensing equipment and challenges in modeling the interaction between the human and exoskeleton (e.g., non-rigid, non-fixed contact points).  Instead, the team develops controllers based on an inverse dynamic model (in the sagittal plane) of only the exoskeleton.

@@

The BLEEX sensitivity amplification control strategy models the torque applied by the human pilot on the exoskeleton as ${\bf d}$ (assuming no outside disturbances).   Neglecting gravity, the exoskeleton angular velocity is modeled as 
\[\vv = \mG \vr + \mS {\bf d} ,\] 
where $\mG$ is the transfer function from actuator inputs ($\mG$ is the exoskeleton's dynamics), $\vr$ is the actuator input, and $\mS$ is the sensitivity or transfer function from human torque to exoskeleton angular velocity.

### 2015-10-01 exos/exosText_MIT.tex
more updates

 The MIT exoskeleton design has a total of fourteen degrees of freedom, three at each hip, one at each knee, two at each ankle, and one in each foot.  A cam mechanism implemented at the hip joint enables hip ab/adduction.  This system was necessary because, during abduction, there is a length difference between the operator and exoskeleton leg which results from having non-collocated centers of rotation.  The cam mechanism automatically adjusts the exoskeleton leg length such that the center of rotation of the exoskeleton hip is projected onto the biological hip center of rotation.  A picture of the MIT exoskeleton is shown in Figure \ref{fig:MITsuit}.
 \begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/MIT/MITsuit}
  \caption{ Sensors placed on the exoskeleton leg \cite{mit_control_2006}.}
%  \vspace{-0.2in}
 \label{fig:MITsuit}   
 \end{figure} 
% 
 The interface between the suit and operator consists of shoulder straps, a compliant waist belt, thigh cuffs, and a shoe connection.  The MIT exoskeleton employs a custom-built sensor that measures the interaction force between the exoskeleton and the operator?s thigh. The sensor consisted of a spring pack, where the deflection of the passive springs was measured with a spring-loaded linear potentiometer.

 The exoskeleton implements a {\bf variable damper knee mechanism} in the flexion/extension degree of freedom using a magnetorheological damper.  The damper can exert a {\bf maximum braking torque of 60 Nm and consums on average 1W of electrical power}.  Additionally, a linear spring at the ankles captures negative energy during dorsilflexion.  The energy stored is released during plantar flexion to add energy to the system as the foot comes off the ground. A urethane spring (247 kN/m) is used as a liner spring in a lever compression assembly.  With a lever arm of 0.0381 m, the assembly yields a spring rotary stiffness of 356 Nm/rad.

@@

The suit uses a 48V battery pack and is instrumented with \textbf{rotary potentiometers} at the hip and knee.  Additionally, \textbf{strain gauges} were placed on the structure of the exoskeleton shank to measured sagittal bending moment and compression force.

@@

 The knee controller in the MIT exoskeleton effectively functions independently from he hip controller.  During knee strike, the variable damper in the knee exerts a torque proportional to the rotational velocity of the knee joint.  A residual magnetic field remains in the knee joint after it is turned off, creating a resistive torque.  The knee therefore needs to be actively demagnetized at full extension during the last phase of stance to allow it to swing freely during the subsequent swing phase.  The state machine diagram that illustrates how the various phase transitions for the knee joint were implemented using the exoskeleton's on-board sensors is in Figure \ref{fig:kneeControl}.
\begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/MIT/kneeControl}
  \caption{State-machine diagram for the knee controller.}
%  \vspace{-0.2in}
 \label{fig:kneeControl} 
 \end{figure}

@@

 The MIT hardware and control system present a well designed system that focuses primarily on efficiency in design.  We generally support the series-elastic linear ball screw hardware design for their mechanical efficiency as well as ability to perform high fidelity force control.  Although, these drives will again be limited by bandwidth issues in highly-dynamic maneuvers.  The use of a strain gauge in the shank was also a novel means of measuring ground reaction forces.  This design may offer benefits in terms of repeatability over different operators, as foot fall patterns may include significant variations for different users or even in the same user in different environments.

>>>>

 The MIT exoskeleton design has a total of fourteen degrees of freedom, three at each hip, one at each knee, two at each ankle, and one in each foot.  A cam mechanism implemented at the hip joint enables hip ab/adduction.  This system was necessary because, during abduction, there is a length difference between the operator and exoskeleton leg which results from having non-collocated centers of rotation.  The cam mechanism automatically adjusts the exoskeleton leg length such that the center of rotation of the exoskeleton hip is projected onto the biological hip center of rotation.  A picture of the MIT exoskeleton is shown in Figure \ref{fig:MITsuit}.
 \begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/MIT/MITsuit}
  \caption{ Sensors placed on the exoskeleton leg \cite{mit_control_2006}.}
%  \vspace{-0.2in}
 \label{fig:MITsuit}   
 \end{figure} 
% 
 The interface between the suit and operator consists of shoulder straps, a compliant waist belt, thigh cuffs, and a shoe connection.  The MIT exoskeleton employs a custom-built sensor that measures the interaction force between the exoskeleton and the operator's thigh. The sensor consisted of a spring pack, where the deflection of the passive springs was measured with a spring-loaded linear potentiometer.

 The exoskeleton implements a {\bf variable damper knee mechanism} in the flexion/extension degree of freedom using a magnetorheological damper.  The damper can exert a {\bf maximum braking torque of 60 Nm and consumes on average 1W of electrical power}.  Additionally, a linear spring at the ankles captures negative energy during dorsilflexion.  The energy stored is released during plantar flexion to add energy to the system as the foot comes off the ground. A urethane spring (247 kN/m) is used as a liner spring in a lever compression assembly.  With a lever arm of 0.0381 m, the assembly yields a spring rotary stiffness of 356 Nm/rad.

@@

The suit uses a 48V battery pack and is instrumented with \textbf{rotary potentiometers} at the hip and knee.  Additionally, \textbf{strain gages} were placed on the structure of the exoskeleton shank to measured sagittal bending moment and compression force.

@@

 The knee controller in the MIT exoskeleton effectively functions independently from the hip controller.  During knee strike, the variable damper in the knee exerts a torque proportional to the rotational velocity of the knee joint.  A residual magnetic field remains in the knee joint after it is turned off, creating a resistive torque.  The knee therefore needs to be actively demagnetized at full extension during the last phase of stance to allow it to swing freely during the subsequent swing phase.  The state machine diagram that illustrates how the various phase transitions for the knee joint were implemented using the exoskeleton's on-board sensors is in Figure \ref{fig:kneeControl}.
\begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/MIT/kneeControl}
  \caption{State-machine diagram for the knee controller.}
%  \vspace{-0.2in}
 \label{fig:kneeControl} 
 \end{figure}

@@

 The MIT hardware and control system present a well designed system that focuses primarily on efficiency in design.  We generally support the series-elastic linear ball screw hardware design for their mechanical efficiency as well as ability to perform high fidelity force control.  Although, these drives will again be limited by bandwidth issues in highly-dynamic maneuvers.  The use of a strain gage in the shank was also a novel means of measuring ground reaction forces.  This design may offer benefits in terms of repeatability over different operators, as foot fall patterns may include significant variations for different users or even in the same user in different environments.

### 2015-10-01 exos/exosText_bodyExt.tex
more updates

 Except for forearm pronosupination, each of the actuators in the Body Extender suit are driven by a linear actuator that is composed of an electrical motor, incremental encoder, ball screw, and angular contact ball bearings.  The motors are frameless DC torque controlled motors.  The {\bf peak torque of the motors is 7 Nm, with a maximum continuous torque of 6 Nm}.  Each motor weighs 1.4 kg.  The {\bf linear actuator assembly is able to supply 8000 N with a total weight of 2.4 kg}.

@@

The feedback linearizing controller in \cite{} operates under two fundamental assumptions and only appears to relate directly to the case of a single manipulator arm.  The first assumption of the controller is that the mass and inertial properties of an object being manipulated are estimated online.  The second is that the forces measured between the operator and suit can be modeled as
\begin{equation}
\vF_s = \mM_p \dot{\vV}_e + \vG_p,
\label{eq:forceEq}
\end{equation}
where $\mM_p$ is an effective mass of the end effector, $\dot{\vV}_e$ is the acceleration of the end effector, and $\mG_p$ is an effective gravitational term.  Note that the gravitational term is zero if the arm is not holding an external load.  The importance of these two assumptions will be more clear after the analytical forms of the feedback linearizing controllers are introduced.

@@

The Body Extender suit is one of the only fully-actuated upper and lower extremity exoskeletons available today.  Clearly, for the task of handling extremely heavy objects, a fully actuated suit may be the best option, but, generally speaking, having the addition weight due to actuating every degree of freedom in the suit is not ideal when considering dynamic operation.

@@

>>>>

 Except for forearm pronosupination, each of the actuators in the Body Extender suit are driven by a linear actuator that is composed of an electrical motor, incremental encoder, ball screw, and angular contact ball bearings.  The motors are frame-less DC torque controlled motors.  The {\bf peak torque of the motors is 7 Nm, with a maximum continuous torque of 6 Nm}.  Each motor weighs 1.4 kg.  The {\bf linear actuator assembly is able to supply 8000 N with a total weight of 2.4 kg}.

@@

The feedback linearizing controller in 
%\cite{}
operates under two fundamental assumptions and only appears to relate directly to the case of a single manipulator arm.  The first assumption of the controller is that the mass and inertial properties of an object being manipulated are estimated online.  The second is that the forces measured between the operator and suit can be modeled as
\begin{equation}
\vF_s = \mM_p \dot{\vV}_e + \vG_p,
\label{eq:forceEq}
\end{equation}
where $\mM_p$ is an effective mass of the end effector, $\dot{\vV}_e$ is the acceleration of the end effector, and $\mG_p$ is an effective gravitational term.  Note that the gravitational term is zero if the arm is not holding an external load.  The importance of these two assumptions will be more clear after the analytical forms of the feedback linearizing controllers are introduced.

@@

The Body Extender suit is one of the only fully-actuated upper and lower extremity exoskeletons available today.  Clearly, for the task of handling extremely heavy objects, a fully actuated suit may be the best option, but, generally speaking, having the additional weight due to actuating every degree of freedom in the suit is not ideal when considering dynamic operation.

@@

The speed of the actuators (approximately 60 deg/s) is far to slow for
dynamic human movement.

### 2015-10-01 exos/exosText_ihmc.tex
more updates

Although the IHMC MAE was primarily designed as a disabled assist exoskeleton, the device is also provides a ``performance enhancement" control mode, which is within the scope of this document.  The main design principle highlighted in the IHMC MAE hardware (a recurring theme throughout exoskeleton literature) is that directly measuring user intent can be difficult for a number of reasons.  The IHMC designers resolve the issue using a new control and feedback system, which for the IHMC MAE, is primarily based on a unique actuation scheme.

@@

The spring mechanism, which sits between the gearbox and joint output, is shown in Figure \ref{fig:IHMCSEA}.  The mechanism uses linear die springs and an angular encoder to measure the joint torque output.  The choice of linear die spring was based on the spring's predictable force to displacement function as well as favorable energy storage to weight characteristics.  
\begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/ihmc/seaAssm}
  \caption{Schematic of the RSEA elastic mechanism design, shown in the zero force position. The encoder reads a scale tape on the curved cylindrical surface. Note that the actuator housing and bearing have been omitted for clarity.}
%  \vspace{-0.2in}
 \label{fig:IHMCSEA}   
 \end{figure}

>>>>

Although the IHMC MAE was primarily designed as a disabled assist exoskeleton, the device also provides a ``performance enhancement" control mode, which is within the scope of this document.  The main design principle highlighted in the IHMC MAE hardware (a recurring theme throughout the exoskeleton literature) is that directly measuring user intent can be difficult for a number of reasons.  The IHMC designers resolve the issue using a new control and feedback system, which for the IHMC MAE, is primarily based on a unique actuation scheme.

@@

The spring mechanism, which sits between the gearbox and joint output, is shown in Figure \ref{fig:IHMCSEA}.  The mechanism uses linear die springs and an angular encoder to measure the joint torque output.  The choice of linear die spring was based on the spring's predictable force to displacement function as well as favorable energy storage to weight characteristics.  
\begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/ihmc/seaAssm}
  \caption{Schematic of the SEA elastic mechanism design, shown in the zero force position. The encoder reads a scale tape on the curved cylindrical surface. Note that the actuator housing and bearing have been omitted for clarity.}
%  \vspace{-0.2in}
 \label{fig:IHMCSEA}   
 \end{figure}

### 2015-10-01 exos/hal.tex
more updates

Human operators attach to HAL at the waist with a belt, and at the calf and thigh using harnesses. HAL's frame does not transfer load to the ground.  Instead, HAL adjusts hip, knee and ankle torques to amplify its operator's torque. The exoskelton's sensory system includes bioelectric sensing (including \textbf{sEMG}), \textbf{angular sensors}, \textbf{acceleration sensors}, and \textbf{center of pressure / center of gravity} (COP/COG) sensors.  COP/COG sensing is provided through shoes with ground reaction force sensors.  The joint measurements are provided by potentiometers.
In at least one version of HAL the bioelectric data comes from sEMG sensors installed below the operator's hip and above the knee (on both front and back).
An IMU installed in HAL's backpack is used to estimate torso pose.

@@

The second control strategy, a model-based approach for human strength augmentation, estimates human intention from sEMG activity and provides power to augment torque provided by the operator.  A relatively autonomous torque assist strategy in \cite{HALmodelControlKnee2010} recognizes user intention to take a step by thresholding sEMG data.  The approach provides a knee torque response including an assitive torque component, a viscous damping component that reduces high velocity motion for safety, and a gravity compensation torque.  
%In \cite{HALvTorqueImp2002}, HAL uses myoelectric data to estimate muscle torque based on the difference between flexor and extensor muscle activities.

>>>>

Human operators attach to HAL at the waist with a belt, and at the calf and thigh using harnesses. HAL's frame does not transfer load to the ground.  Instead, HAL adjusts hip, knee and ankle torques to amplify its operator's torque. The exoskeleton's sensory system includes bioelectric sensing (including \textbf{sEMG}), \textbf{angular sensors}, \textbf{acceleration sensors}, and \textbf{center of pressure / center of gravity} (COP/COG) sensors.  COP/COG sensing is provided through shoes with ground reaction force sensors.  The joint measurements are provided by potentiometers.
In at least one version of HAL the bioelectric data comes from sEMG sensors installed below the operator's hip and above the knee (on both front and back).
An IMU installed in HAL's backpack is used to estimate torso pose.

@@

The second control strategy, a model-based approach for human strength augmentation, estimates human intention from sEMG activity and provides power to augment torque provided by the operator.  A relatively autonomous torque assist strategy in \cite{HALmodelControlKnee2010} recognizes user intention to take a step by thresholding sEMG data.  The approach provides a knee torque response including an assistive torque component, a viscous damping component that reduces high velocity motion for safety, and a gravity compensation torque.  
%In \cite{HALvTorqueImp2002}, HAL uses myoelectric data to estimate muscle torque based on the difference between flexor and extensor muscle activities.

### 2015-10-01 exos/xor.tex
more updates

The XoR weighs 30kg and includes 10 DOF with 6 active joints (flexion / extension of hip, knee, and ankles) and 6 passive (hip abduction/adduction joints, hip rotation, and ankle adduction/abduction).  The active joints are powered by \textbf{hybrid actuators} comprised of an \textbf{air muscle} and an \textbf{electric motor}.  The hybrid actuation scheme uses a unilateral air muscle layout to compensate for gravity and bilateral electric motors with relatively small gear ratio (57.5) to serve as the dynamic compensator.

The actuation scheme is complementary in that the electric motors have a quick response time and produce high peak torque for short amounts of time, while air muscles have a delayed response (true of pneumatic systems in general due to the compressibility of air) with better power density and sustained torque.  The hybrid drive system sums the two to develop a desired torque profile that achieves the benefits of both and reduces weight and motor size.  The system offers high torque control with negligable stick-slip and backdrivability.

@@

To avoid relying on high gain feedback, XoR incorporates user intent through EMG data.  The controller rectifies and low pass filters (10 Hz cut-off) data measured in the pilot's quadriceps femoris, tensor fasciae latate, gluteus medius, and tibialis anterior. 
It sets desired joint angles and velocities, ${\bf x} = ({\bf \theta}, \dot{\bf \theta})$, from the EMG data, ${\bf u} = (EMG_1, \dots , EMG_n)$, using a linear prediction model, 
\[{\bf x}(k+1) = \mA{\bf x}(k)+\mB{\bf u}(k),\].  
The controller uses the angular measurements required to perform CCA and EMG data to derive the model's $\mA$ and $\mB$ terms.  The predicted state from EMG model provides the desired input to yet another PD controller: 
\[\tau_i = K_p (\theta_i^d - \theta_i) + K_d (\dot\theta_i^d - \dot \theta_i),\] 
with gains of $K_p = 1000$ and $K_d = 100$.

In hip tracking experiments, the team found EMG data to be helpful.  When using EMG signals, XoR achieved a mean squared error of $1.8E^{-3}$.  Without the EMG data, the mean squared error increased to $2.1$.  Additional experiments confirm the air muscles can effectively perform gravity compensation when used in conjunction with electric actuators (to correct for torque errors induced by inaccuracies in the model mapping position to toque in air muscles).

>>>>

The XoR weighs 30kg and includes 10 DOF with 6 active joints (flexion / extension of hip, knee, and ankles) and 6 passive (hip abduction/adduction joints, hip rotation, and ankle adduction/abduction).  The active joints are powered by \textbf{hybrid actuators} comprised of an \textbf{air muscle} and an \textbf{electric motor}.  The hybrid actuation scheme uses a unilateral air muscle layout to compensate for gravity and bilateral electric motors with a relatively small gear ratio (57.5) to serve as the dynamic compensator.

The actuation scheme is complementary in that the electric motors have a quick response time and produce high peak torque for short amounts of time, while air muscles have a delayed response (true of pneumatic systems in general due to the compressibility of air) with better power density and sustained torque.  The hybrid drive system sums the two to develop a desired torque profile that achieves the benefits of both and reduces weight and motor size.  The system offers high torque control with negligible stick-slip and backdrivability.

@@

To avoid relying on high gain feedback, XoR incorporates user intent through EMG data.  The controller rectifies and low pass filters (10 Hz cut-off) data measured in the pilot's quadriceps femoris, tensor fasciae latate, gluteus medius, and tibialis anterior. 
It sets desired joint angles and velocities, ${\bf x} = ({\bf \theta}, \dot{\bf \theta})$, from the EMG data, ${\bf u} = (EMG_1, \dots , EMG_n)$, using a linear prediction model, 
\[{\bf x}(k+1) = \mA{\bf x}(k)+\mB{\bf u}(k),\]  
The controller uses the angular measurements required to perform CCA and EMG data to derive the model's $\mA$ and $\mB$ terms.  The predicted state from EMG model provides the desired input to yet another PD controller: 
\[\tau_i = K_p (\theta_i^d - \theta_i) + K_d (\dot\theta_i^d - \dot \theta_i),\] 
with gains of $K_p = 1000$ and $K_d = 100$.

In hip tracking experiments, the team found EMG data to be helpful.  When using EMG signals, XoR achieved a mean squared error of $1.8E^{-3}$.  Without the EMG data, the mean squared error increased to $2.1$.  Additional experiments confirm the air muscles can effectively perform gravity compensation when used in conjunction with electric actuators (to correct for torque errors induced by inaccuracies in the model mapping position to torque in air muscles).

### 2015-10-01 survey.tex
more updates

\title{A Survey of Current Exoskeletons\\ and Their Control Architectures and
Algorithms\\
(Draft 3.2)}

@@

This paper surveys the following exoskeletons in detail:
\begin{itemize}
\item
{\bf BLEEX:} a ``minimal-sensing'' and aggressive model-based control design that has influenced many subsequent exoskeletons.
\item
{\bf HAL:} an EMG controlled exoskeleton, emphasizing recognizing operator intent.
\item
{\bf XoR:} a hybrid Pneumatic/Electric-motor system, emphasizing the
value of using multiple actuator types.
\item
{\bf Body Extender:} a fully actuated full body exoskeleton. This is the
Iron Man option.
\item
{\bf HLEE:} a nice and well presented example of a torque control approach.
\item
{\bf IHMC MAE:} an SEA-based torque controlled exoskeleton, emphasizing
the value of mixing passive mechanisms (local energy storage) and active
actuation.
\item
{\bf MIT Exoskeleton:} an attempt to take advantage of the passive dynamics
of walking.
\item
{\bf RoboKnee:} an example of a single-joint exoskeleton. Powerful single-joint
or minimal exoskeletons are one way to enable operators to move fast
and naturally, and greatly increase their strength.
\end{itemize}

@@

To this end, the hardware systems that we chose to include summaries of in this document were the BLEEX, HAL, XoR, Body Extender, Hydraulic Lower Extremety, IHMC Mobility Assist, MIT, and RoboKnee exoskeletons.  The different systems cover many of the possible choices of actuation strategies, ranging from electric motors, to hydraulics, to pneumatics, and combinations thereof.  Several of the systems used novel transmission strategies as well, focusing on both efficient as well as high-fidelity force control implementations.  The transmission strategies included standard hydraulic systems, hydraulic systems with actively controlled bypass valves, ball-screw mechanisms that include series elastic components, as well as pneumatically actuated pulley-tendon and belt-drive mechanisms.

@@

Directly measuring user intent is difficult to implement using standard sensing modalities.  For instance, EMG data is notoriously noisy and presents both modeling and calibration challenges.  It would also be extremely difficult to keep these sensors in place to obtain accurate readings in dynamic environments or combat scenarios.  For similar reasons, many systems avoid the use of load cells or complicated sensor networks to accurately obtain operator-robot interaction forces / torques.

@@

% Based on these observations, our main recommendation is that a \emph{modified get out of the way} control strategy provides a reasonable policy for dynamic exoskeletons.
Based on these observations, we recommend (at least in the short term) fusing the two predominant approaches to exoskeleton control. That is, combining sensitivity amplification strategies with direct user measurement to robustly distinguish user intent from external disturbances.
% For instance, the get out of the way controller used in the Hydraulic Lower Extremity Exoskeleton would provide a starting point for this strategy.  This controller utilizes IMUs, foot contact sensors, and joint torque sensors to implement a user-reactive gravity compensating strategy that does not directly measure the user.  Augmenting the sensor suite with sensors that that measure user generated forces could allow this system to identify and reject external disturbances while simultaneously shadowing user motion.
%
% We emphasize that user intent is critical, and current sensing challenges must be addressed in order to develop dynamic exoskeletons capable of serving in combat scenarios.  As a stop-gap, it may be possible to combine low information density EMG (thresholded on/off muscle activation signals) and human force/torque or angular measurements to filter user intent.  Such an approach may prove useful in a hybrid strategy, filtering external disturbance forces during sensitivity amplification control.  
% As a longer term solution, we highlight research in the area of MEMs and nano-fabrication, which has lead to the development of new, flexible ``robotic skin'' sensing arrays.  Integrating these sensing arrays into operator clothing (or attaching directly to the exoskeleton), it may be possible to directly measure human interaction forces (i.e., estimating pressure / force and exoskeleton contact locations), which was not feasible in the recent past.  Dense sensing arrays add redundancy and can improve measurement reliability. We recommend further study and the development of more advanced sensory systems to address the challenges inherent in dynamic exoskeleton control.
%
For instance, it may be possible to use low information density EMG (thresholded on/off muscle activation signals) with robot force/torque and angular measurements to filter disturbances from operator-robot interaction forces.  As a longer term solution, we highlight research in the area of MEMs, nano-fabrication, soft sensors, etc., which has lead to the development of new, flexible ``robotic skin'' sensing arrays.  Integrating these sensing arrays into operator clothing (or attaching directly to the exoskeleton), it may be possible to directly measure human interaction forces (i.e., estimating pressure / force and exoskeleton contact locations), which was not feasible in the recent past.  Dense sensing arrays add redundancy and can improve measurement reliability and, thereby, feedback control performance.

@@

Lastly, the importance of producing hardware that is physically capable of achieving dynamic performance is fundamentally linked to how mass is distributed throughout the system.  In general, it is desirable to minimize mass, particularly in distal regions. The design objective is often achieved by focusing actuation and leaving certain degrees of freedom passive (leading to an underactued system).  Underactuted control is a complicated problem that is often ignored in the sense that controllers assume the operator will provide necessary forces and torques for the un-actuated degrees of freedom.  Though it is possible to deal with the underactuated control problem directly (e.g., optimization-based controllers), there is a balance between feedback and feedforward control.  
% The better one is, the less reliant the system is on the other.  
By developing better sensors and feedback controllers that directly measure and shadow operator intent, exoskeleton controllers will become more robust and less sensitive to errors due to underactuated system assumptions and dynamic model inaccuracies.

>>>>

\title{A Survey of Current Exoskeletons\\ and Their Control Architectures and
Algorithms\\
(Draft 4.0)}

@@

This paper surveys the following exoskeletons in detail:
\begin{itemize}
\item
{\bf BLEEX:} a ``minimal-sensing'' and aggressive model-based control design that has influenced many subsequent exoskeletons.
\item
{\bf HAL:} an EMG controlled exoskeleton, emphasizing recognizing operator intent.
\item
{\bf XoR:} a hybrid pneumatic/electric-motor system, emphasizing the
value of using multiple actuator types. It also combines the use
of brain signals and EMG for estimating operator intent.
\item
{\bf Body Extender:} a fully actuated full body exoskeleton. This is the
Iron Man option.
\item
{\bf HLEE:} a nice and well presented example of a torque control approach.
\item
{\bf IHMC MAE:} an SEA-based torque controlled exoskeleton, emphasizing
the value of mixing passive mechanisms (local energy storage) and active
actuation.
\item
{\bf MIT Exoskeleton:} an attempt to take advantage of the passive dynamics
of walking.
\item
{\bf RoboKnee:} an example of a single-joint exoskeleton. Powerful single-joint
or minimal exoskeletons are one way to enable operators to move fast
and naturally, and greatly increase their strength.
\end{itemize}

@@

To this end, the hardware systems that we chose to include summaries of in this document were the BLEEX, HAL, XoR, Body Extender, Hydraulic Lower Extremity, IHMC Mobility Assist, MIT, and RoboKnee exoskeletons.  The different systems cover many of the possible choices of actuation strategies, ranging from electric motors, to hydraulics, to pneumatics, and combinations thereof.  Several of the systems used novel transmission strategies as well, focusing on both efficient as well as high-fidelity force control implementations.  The transmission strategies included standard hydraulic systems, hydraulic systems with actively controlled bypass valves, ball-screw mechanisms that include series elastic components, as well as pneumatically actuated pulley-tendon and belt-drive mechanisms.

@@

Directly measuring user intent is difficult to implement using standard sensing modalities.  For instance, EMG data is notoriously noisy and presents both modeling and calibration challenges.  It would also be extremely difficult to keep these sensors in place to obtain accurate readings in dynamic environments or activities.  For similar reasons, many systems avoid the use of load cells or complicated sensor networks to accurately obtain operator-robot interaction forces/torques.

@@

% Based on these observations, our main recommendation is that a \emph{modified get out of the way} control strategy provides a reasonable policy for dynamic exoskeletons.
Based on these observations, we recommend (at least in the short term) fusing the two predominant approaches to exoskeleton control. That is, combining sensitivity amplification strategies with direct user measurement to robustly distinguish user intent from external disturbances.
% For instance, the get out of the way controller used in the Hydraulic Lower Extremity Exoskeleton would provide a starting point for this strategy.  This controller utilizes IMUs, foot contact sensors, and joint torque sensors to implement a user-reactive gravity compensating strategy that does not directly measure the user.  Augmenting the sensor suite with sensors that that measure user generated forces could allow this system to identify and reject external disturbances while simultaneously shadowing user motion.
%
% We emphasize that user intent is critical, and current sensing challenges must be addressed in order to develop dynamic exoskeletons capable of serving in combat scenarios.  As a stop-gap, it may be possible to combine low information density EMG (thresholded on/off muscle activation signals) and human force/torque or angular measurements to filter user intent.  Such an approach may prove useful in a hybrid strategy, filtering external disturbance forces during sensitivity amplification control.  
% As a longer term solution, we highlight research in the area of MEMs and nano-fabrication, which has led to the development of new, flexible ``robotic skin'' sensing arrays.  Integrating these sensing arrays into operator clothing (or attaching directly to the exoskeleton), it may be possible to directly measure human interaction forces (i.e., estimating pressure/force and exoskeleton contact locations), which was not feasible in the recent past.  Dense sensing arrays add redundancy and can improve measurement reliability. We recommend further study and the development of more advanced sensory systems to address the challenges inherent in dynamic exoskeleton control.
%
For instance, it may be possible to use low information density EMG (thresholded on/off muscle activation signals) with robot force/torque and angular measurements to filter disturbances from operator-robot interaction forces.  As a longer term solution, we highlight research in the area of MEMs, nano-fabrication, soft sensors, etc., which has lead to the development of new, flexible ``robotic skin'' sensing arrays.  Integrating these sensing arrays into operator clothing (or attaching directly to the exoskeleton), it may be possible to directly measure human interaction forces (i.e., estimating pressure / force and exoskeleton contact locations), which was not feasible in the recent past.  Dense sensing arrays add redundancy and can improve measurement reliability and, thereby, feedback control performance.

@@

Lastly, the importance of producing hardware that is physically capable of achieving dynamic performance is fundamentally linked to how mass is distributed throughout the system.  In general, it is desirable to minimize mass, particularly in distal regions. The design objective is often achieved by focusing actuation and leaving certain degrees of freedom passive (leading to an underactuated system).  Underactuated control is a complicated problem that is often ignored in the sense that controllers assume the operator will provide necessary forces and torques for the un-actuated degrees of freedom.  Though it is possible to deal with the underactuated control problem directly (e.g., optimization-based controllers), there is a balance between feedback and feedforward control.  
% The better one is, the less reliant the system is on the other.  
By developing better sensors and feedback controllers that directly measure and shadow operator intent, exoskeleton controllers will become more robust and less sensitive to errors due to underactuated system assumptions and dynamic model inaccuracies.
